Squintability and Other Metrics for Assessing Projection Pursuit Indexes, and Guiding Optimization Choices

### 2025-03-04 jso.tex
proofread done

The primary goal of the study reported here is to investigate the
performance of JSO in PP optimization for the guided tour. It is of
interest to assess how quickly and closely the optimizer reaches a
global optimum, for various PP indexes that may have differing
complexities. To observe the performance of JSO with different types of
PP indexes, metrics are introduced to capture specific properties of the
index including squintability (based on Tukey and Tukey 1981's squint
angle) and smoothness. We mathematically define metrics for
squintability and smoothness, which is a new contribution for PP
research. A series of simulation experiments are conducted using JSO to
detect different target patterns (pipe and sine-wave) with different PP
indexes (holes, MIC, TIC, dcor, loess, splines, skinny, and stringy), as
well as JSO's hyper-parameter choices (number of jellyfish and maximum
number of tries). This work should facilitate better optimization for PP
and guide the choice of optimizer when designing new PP indexes.

@@

To solve the optimization problem embedded in the PPGT with JSO, a
starting projection, an index function, the number of jellyfish, and the
maximum number of trials (tries) are provided as input. Then, the
current projection is evaluated by the index function. The projection is
then moved in a direction determined by a random factor, influenced by
how far along we are in the optimization process. Occasionally,
completely new directions may be taken like a jellyfish might with ocean
currents. A new projection is accepted if it is an improvement compared
to the current one, rejected otherwise. This process continues and
iteratively improves the projection, until the pre-specified maximum
number of trials is reached.

\begin{tcolorbox}[enhanced jigsaw, bottomrule=.15mm, leftrule=.75mm, breakable, coltitle=black, colback=white, colbacktitle=quarto-callout-note-color!10!white, arc=.35mm, colframe=quarto-callout-note-color-frame, toptitle=1mm, opacitybacktitle=0.6, opacityback=0, left=2mm, bottomtitle=1mm, rightrule=.15mm, toprule=.15mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, titlerule=0mm]

@@

This set of simulations is expanded to assess the performance of JSO
relative to the index properties, with a second data set type and
additional PP indexes. The second data set has a sine wave in two
dimensions and Gaussian noise in the remaining dimensions. Seven
additional PP indexes that should be able to extract the sine wave are
used: \texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
\texttt{splines2d}, \texttt{skinny} and \texttt{stringy2}, following Laa
and Cook (2020). (The supplementary material includes definitions of
these PP indexes, and descriptions of the pipe and sine wave data
simulation.). The JSO hyper-parameters, number of jellies and maximum
tries, are also included. This results in a total of 76 scenarios,
comprising of 30 computed on the pipe data and 46 on the sine wave data.
The choice of 4-12 dimensions covers the range of easy to relatively
difficult for the optimizers to detect the structured projections. The
same approach is used, that each experiment is repeated 50 times and
success rate is calculated at each level of the simulation design.

@@

Recall that smoothness and squintability are metrics used to
characterize the index function for PP tasks and are invariant to the
optimizer and its hyper-parameters. Thus simulations with different JSO
hyper-parameters for the same PP task share the same smoothness value.
In total, smoothness and squintability metrics are calculated on the 22
PP tasks in our simulation.

@@

A generalized linear model is fitted using a quasibinomial family and a
logit link function to assess the factors affecting the success rate.
Predictors are smoothness, squintability, data dimension, and JSO
hyper-parameters.

@@

\caption{\label{fig-proj}Visual comparison of JSO and CRS results, using
optimal data projections obtained over 50 simulations of the pipe data.
Rows correspond to data dimension. Columns correspond to quantiles of
the index values, with 0 being minimum, 100 being maximum, and 50 the
median. JSO achieves the better views of the pipe generally than CRS. As
dimension increases both have more difficulty finding the pipe.}

@@

The performance of JSO is assessed relative to the existing CRS
optimizer based on the optimal projection obtained. The final
projections for the pipe data sets found by the two optimizers are shown
in Figure~\ref{fig-proj}, with rows corresponding to data dimension. The
50 simulations in each data dimension are sorted by index value, and the
projections corresponding to each 10th quantile value are shown, with
minimum at left and maximum at right. Index value is printed on the
plot. The purpose is to summarize the views of the data resulting from
different optimization, and hence compare results between the two
optimizers. Generally, the JSO does a more consistent job of finding the
pipe structure clearly. As the data dimension increases, both optimizers
struggle and less clearly capture the circle shape.

@@

The effect of JSO hyper-parameters (number of jellyfish and the maximum
number of tries) on the success rate is summarized in
Figure~\ref{fig-proportion}. The uncertainty is quantified through 500
bootstrap replicates for each case. As the number of jellyfish and
maximum tries increase, the success rate also increases. For problems
with relatively lower dimensional (4, 6) search spaces, small parameter
values (20 jellyfish and a maximum of 50 tries) are sufficient to
achieve a high success rate. Many more jellyfish (100 jellyfish and a
maximum of 100 tries) are needed to obtain a mediocre success rate for
dimensions 10 and 12. This suggests that even JSO will struggle with
high dimensions, and require substantial more computational time to have
an acceptable success rate, especially for difficult PP indexes based on
scagnostics.

@@

\caption{\label{fig-proportion}Summary of the relationship between JSO
hyper-parameters on optimization success, using the holes index for pipe
data of dimensions 4-12. Bootstrap samples show the variability in
success rate. Success rate mostly plateaus by 50 jellies. The JSO has
some difficulty finding the pipe when dimension is higher than 8!
Maximum number of tries has little effect.}

@@

Smoothness and squintability are calculated across a collection of pipe
and sine wave finding problems to construct the relationship between
success rate, JSO hyper-parameters, and index properties.
Table~\ref{tbl-smoothness-squintability} presents values computed for a
subset of PP indexes on the sine wave data for dimensions 4-8. (A full
table of the parameters estimated from the Gaussian process (outputscale
\(\eta\), lengthscale \(\ell\), smoothness \(\nu\), and nugget
\(\sigma\)) and from the scaled logistic function (\(\theta_1\) to
\(\theta_4\)) for calculating smoothness and squintability for all PP
indexes and data combinations is included in the supplementary
materials). The column \(\nu\) is used as the smoothness metric and the
column \(\varsigma\) is calculated as equation
(\ref{eq-squintability-parametric}) as the squintability metric.

Figure~\ref{fig-idx-proj-dist} shows a sample of traces from the indexes
in Table~\ref{tbl-smoothness-squintability} for the 6D sine wave data.
The traces the PP index values computed for projections along an
interpolated path between a noise projection and the optimal projection
(sine wave). (Note index values were self-standardized to range between
0 and 1, for comparison.) The \texttt{splines} index is very smooth with
a gradual increase in value between the noise projection and the optimal
projection. The \texttt{TIC} index is slightly less smooth and slightly
less gradual change in value. Least desirable is the \texttt{skinny}
index which is noisier and the optimization needs to be much closer to
the optimal angle to have a higher value. This illustrates how the new
measures for squintability and smoothness work to describe the PP
indexes.

@@

\begin{tabular}{|llr|rr|llr|rr|llr|rr|llr|rr|llr|rr}
\toprule
shape & index & d & $\nu$ & $\varsigma$\\
\midrule
sine & splines & 4 & 3.5489 & 0.5723\\
sine & splines & 6 & 3.1149 & 0.5415\\
sine & splines & 8 & 2.7400 & 0.5394\\
sine & TIC & 4 & 3.4854 & 0.3839\\
sine & TIC & 8 & 2.7223 & 0.3785\\
sine & TIC & 6 & 3.0807 & 0.3747\\
sine & skinny & 6 & 2.2352 & 0.0973\\
\bottomrule
\end{tabular}

@@

Table~\ref{tbl-mod-output} presents the results from fitting a logistic
regression model using the proportion of success as the response and
smoothness, squintability, dimension, number of jellyfish and maximum
number of tries, as predictors. The fit suggests that the JSO success
rate is only affected by squintability, dimension and number of jellies.
As expected, the JSO success rate is higher with higher squintability
and/or more jellyfish, and lower when dimension is higher
(higher-dimensional optimization is more difficult). Interestingly, the
JSO is unaffected by the smoothness of the index function. This is
consistent with the way random search algorithms jump from value to
value without taking local regularity into account, as opposed to
gradient-based optimizers for example. Allowing for more tries also does
not affect success rate significantly. A unit increase in squintability
increases the success rate by 24\%. As dimension increases by one the
success rate almost halves. Increasing the number of jellies by 10
increases the success rate by 32\%. Table~\ref{tbl-joint-output} shows
the fit statistics supporting that the model is reasonably strong.

Note that the addition of interaction terms for squintability and data
dimension, and smoothness and data dimension, improves the model a small
amount, but does not change the interpretation above.

@@

\caption{\label{tbl-mod-output}Jellyfish success rate relative to index
properties and jellyfish hyper-parameters. This is the summary from a
logistic regression fit to smoothness, squintability, dimension, number
of jellyfish and maximum number of tries. Interestingly, squintability
and dimension strongly affect jellyfish optimization success. The number
of jellies marginally affects success, but index smoothness, and
increasing the number of tries do not.}

@@

\toprule
\textbf{Characteristic} & \textbf{OR} \textbf{(95\% CI)}\textsuperscript{\textit{1}} & \textbf{p-value} \\ 
\midrule\addlinespace[2.5pt]
Smoothness & 1.00 (0.90 to 1.11) & 0.928 \\ 
Squintability & 1.24 (1.13 to 1.39) & {\bfseries <0.001} \\ 
Dimension & 0.57 (0.44 to 0.73) & {\bfseries <0.001} \\ 
Number of jellyfish & 1.32 (1.17 to 1.50) & {\bfseries <0.001} \\ 
Maximum number of tries & 1.06 (0.94 to 1.21) & 0.333 \\ 
\bottomrule

@@

\\
\toprule
\textbf{Deviance} & \textbf{DF Residual} & \textbf{Null Deviance} & \textbf{DF Null}\\
\midrule
20.69 & 70 & 49.29 & 75\\
\bottomrule

@@

This study intends to serve as an illustration that the proposed
measures effectively captures the key factors influencing the success
rate of PP.

@@

The simulation studies have compared the JSO with CRS and shown how
smoothness and squintability affect the JSO success rate. Here we
explain how they can be used for new index and optimizer development.
Using the JSO optimizer for PPGT requires a change in user behavior. For
all the existing optimizers a single optimization path is followed. The
JSO has multiple optimization paths, and needs additional tools to
incorporate into the PPGT, as explained here.

@@

To use JSO for PPGT in the \texttt{tourr} package, specify
\texttt{search\_f\ =\ search\_jellyfish} in the guided tour. Unlike
existing optimizers, the animation function \texttt{animate\_*()} won't
directly render the tour path for JSO due to the generation of multiple
tour paths. To visualize the path visited by each individual jellyfish,
assign the animation to an object
(\texttt{res\ \textless{}-\ animate\_xy(...)}). This will save the bases
visited by JSO, along with relevant metadata, as a tibble data object
(see Zhang et al. 2021 for more details on the data object). The bases
visited for individual jellyfish can then be extracted and viewed using
the \texttt{planned\_tour()} function.

@@

The \texttt{ferrn} package (Zhang et al. 2021) provides functionality
for computing the smoothness and squintability metrics. Both metrics
require sampling random bases using the \texttt{sample\_bases()}
function, followed by the metric calculation with
\texttt{calc\_smoothness()} or \texttt{calc\_squintability()}.

@@

To sample bases for calculating smoothness, the following arguments are
required: the index function, the dataset, and the number of random
bases to sample. The output of \texttt{sample\_bases()} is a data frame
with a list-column of sampled basis matrix and index value.
Particularization is available to speed up the index value computation
through the \texttt{parallel\ =\ TRUE} argument.

The \texttt{calc\_smoothness()} function takes the output from
\texttt{sample\_bases()} and fits a Gaussian process model to the index
values against the sampled basis, as the location, to obtain the
smoothness metric. Starting parameters and additional Gaussian process
arguments can be specified and details of the fit can be accessed
through the \texttt{fit\_res} attribute of the output.

@@

Bases sampling for squintability includes an additional step of
interpolating between the sampled bases and the best basis. This step is
performed when the arguments \texttt{step\_size} and
\texttt{min\_proj\_dist} in the \texttt{sample\_bases()} function are
set to non-NA numerical values. Given the projection distance typically
ranging from 0 to 2, it is recommended to set \texttt{step\_size} to 0.1
or lower, and \texttt{min\_proj\_dist} to be at least 0.5 to ensure a
meaningful interpolation length.

The \texttt{calc\_squintability()} function computes squintability using
two methods: 1) parametrically, by fitting a scaled sigmoid function
through non-linear least square (\texttt{method\ =\ "nls"}), and 2)
non-parametrically, using kernel smoothing (\texttt{method\ =\ "ks"}). A
\texttt{bin\_width} argument is required to average the index values
over the projection distance before the fit. For the parametric case,
the output provides the estimated parameters for the scaled sigmoid
function (\(\theta_1\) to \(\theta_4\)) and the calculated squintability
metric as Equation (\ref{eq-squintability-parametric}). For the
non-parametric case, it shows the maximum gradient attained
(\texttt{max\_d}), the corresponding projection distance
(\texttt{max\_dist}), and the squintability metric as their products.

>>>>

The primary goal of the study reported here is to investigate the
performance of JSO in PP optimization for the guided tour. It is of
interest to assess how quickly and closely the optimizer reaches a
global optimum, for various PP indexes that may have differing
complexities. To observe the performance of JSO with different types of
PP indexes, metrics are introduced to capture specific properties of the
index including squintability (based on Tukey and Tukey 1981's squint
angle) and smoothness. We mathematically define metrics for
squintability and smoothness, which is a new contribution for PP
research. A series of simulation experiments are conducted using JSO to
detect different target patterns (pipe and sine-wave) with different PP
indexes (holes, MIC, TIC, dcor, loess, splines, skinny, and stringy), as
well as JSO's hyper-parameter choices (number of jellyfish and maximum
number of iterations). This work should facilitate better optimization
for PP and guide the choice of optimizer when designing new PP indexes.

@@

To solve the optimization problem embedded in the PPGT with JSO, a
starting projection, an index function, the number of jellyfish, and the
maximum number of iterations are provided as input. Then, the current
projection is evaluated by the index function. The projection is then
moved in a direction determined by a random factor, influenced by how
far along we are in the optimization process. Occasionally, completely
new directions may be taken like a jellyfish might with ocean currents.
A new projection is accepted if it is an improvement compared to the
current one, rejected otherwise. This process continues and iteratively
improves the projection, until the pre-specified maximum number of
trials is reached.

\begin{tcolorbox}[enhanced jigsaw, arc=.35mm, colback=white, coltitle=black, colbacktitle=quarto-callout-note-color!10!white, bottomtitle=1mm, titlerule=0mm, title={Algorithm: Jellyfish Optimizer Pseudo Code}, rightrule=.15mm, breakable, toprule=.15mm, leftrule=.75mm, opacityback=0, opacitybacktitle=0.6, toptitle=1mm, colframe=quarto-callout-note-color-frame, bottomrule=.15mm, left=2mm]

@@

This set of simulations is expanded to assess the performance of JSO
relative to the index properties, with a second data set type and
additional PP indexes. The second data set has a sine wave in two
dimensions and Gaussian noise in the remaining dimensions. Seven
additional PP indexes that should be able to extract the sine wave are
used: \texttt{dcor2d\_2}, \texttt{loess2d}, \texttt{MIC}, \texttt{TIC},
\texttt{splines2d}, \texttt{skinny} and \texttt{stringy2}, following Laa
and Cook (2020). (The supplementary material includes definitions of
these PP indexes, and descriptions of the pipe and sine wave data
simulation.). The JSO hyper-parameters, number of jellyfish and maximum
iterations, are also included. This results in a total of 77 scenarios,
comprising of 30 computed on the pipe data and 47 on the sine wave data.
The choice of 4-12 dimensions covers the range of easy to relatively
difficult for the optimizers to detect the structured projections. The
same approach is used, that each experiment is repeated 50 times and
success rate is calculated at each level of the simulation design.

@@

Recall that smoothness and squintability are metrics used to
characterize the index function for PP tasks and are invariant to the
optimizer and its hyper-parameters. Thus simulations with different JSO
hyper-parameters for the same PP task share the same smoothness value.
In total, smoothness and squintability metrics are calculated on the 23
PP tasks in our simulation.

@@

A generalized linear model is fitted using a quasibinomial family and a
logit link function to assess the factors affecting the success rate.
Predictors are smoothness, squintability, data dimension, and JSO
hyper-parameters. Due to the scale of the smoothness and squintability
measure, the rank of the two metrics, rather than the nominal values,
are used in the model.

@@

\caption{\label{fig-proj}Visual comparison of JSO and CRS results, using
optimal data projections obtained over 50 simulations of the pipe data.
Rows correspond to data dimension. Columns correspond to quantiles of
the index values, with 0 being the minimum, 100 being the maximum, and
50 the median. JSO achieves the better views of the pipe generally than
CRS. As dimension increases both have more difficulty finding the pipe.}

@@

The optimal projection is compared between JSO and CRS for the pipe data
at dimensions 6, 8, 10, and 12 in Figure~\ref{fig-proj}. The columns
show the quantile of index value in the 50 optimal projection, with 0
being the minimum, 100 being the maximum, and 50 the median. The purpose
is to summarize the views of the data resulting from different
optimization, and hence compare results between the two optimizers.
Generally, the JSO does a more consistent job of finding the pipe
structure clearly. As the data dimension increases, both optimizers
struggle and less clearly capture the circle shape.

@@

The effect of JSO hyper-parameters (number of jellyfish and the maximum
number of iteration) on the success rate is summarized in
Figure~\ref{fig-proportion}. The uncertainty is quantified through 500
bootstrap replicates sampled from the simulation for each case. As the
number of jellyfish and maximum iteration allowed increase, the success
rate also increases. For problems in lower dimensional (4, 6) search
spaces, small parameter values (20 jellyfish and 50 iterations) are
sufficient to achieve a high success rate. Many more jellyfish (100
jellyfish and 100 iterations) are neede for finding the pipe structure
in d = 10, and 12. This suggests that even JSO will struggle with high
dimensions, and require substantial more computational time to have an
acceptable success rate, especially for difficult PP indexes based on
scagnostics.

@@

\caption{\label{fig-proportion}Summary of the relationship between JSO
hyper-parameters on optimization success rate, using the holes index for
pipe data of dimensions 4-12. Bootstrap samples show the variability in
success rate. Success rate mostly plateaus by 50 jellyfish. The JSO has
some difficulty finding the pipe when dimension is higher than 8 and
maximum number of iteration has little effect.}

@@

Table~\ref{tbl-smoothness-squintability} presents values computed for a
subset of PP indexes on the sine wave data for dimensions 4-8. (A full
table of the parameters estimated from the Gaussian process (outputscale
\(\eta\), lengthscale \(\ell\), smoothness \(\nu\), and nugget
\(\sigma\)) and from the scaled logistic function (\(\theta_1\) to
\(\theta_4\)) for calculating smoothness and squintability for all PP
indexes and data combinations is included in the supplementary
materials). The column \(\nu\) is used as the smoothness metric and the
column \(\varsigma\) is calculated as equation
(\ref{eq-squintability-parametric}) as the squintability metric.

Figure~\ref{fig-idx-proj-dist} shows a sample of traces from the indexes
in Table~\ref{tbl-smoothness-squintability} for the 6D sine wave data.
This traces the PP index values computed for projections along an
interpolated path between a randomly generated noise projection and the
optimal projection (sine wave). (Note index values were
self-standardized to range between 0 and 1, for comparison.) The
\texttt{splines} index is very smooth with a gradual increase in value
between the noise projection and the optimal projection. The
\texttt{TIC} index is slightly less smooth and slightly less gradual
change in value. Least desirable is the \texttt{skinny} index which is
noisier and the optimization needs to be much closer to the optimal
angle to have a higher value. This illustrates how the new measures for
squintability and smoothness work to describe the PP indexes.

@@

\begin{tabular}{|llr|rr|llr|rr|llr|rr|llr|rr|llr|rr}
\toprule
shape & index & d & $\nu$ & $\varsigma$\\
\midrule
sine & splines & 4 & 3.5489 & 0.5723\\
sine & splines & 6 & 3.1149 & 0.5415\\
sine & splines & 8 & 2.7400 & 0.5394\\
sine & TIC & 4 & 3.4854 & 0.3839\\
sine & TIC & 8 & 2.7223 & 0.3785\\
sine & TIC & 6 & 3.0807 & 0.3747\\
sine & skinny & 6 & 2.2352 & 0.0973\\
sine & skinny & 4 & 2.5213 & 0.0810\\
\bottomrule
\end{tabular}

@@

Table~\ref{tbl-mod-output} presents the results from fitting a logistic
regression model using the proportion of success as the response and
smoothness, squintability, dimension, number of jellyfish and maximum
number of iteration, as predictors. The fit suggests that the JSO
success rate is only affected by squintability, dimension and number of
jellyfish. As expected, the JSO success rate is higher with higher
squintability and/or more jellyfish, and lower when dimension is higher
(higher-dimensional optimization is more difficult). Interestingly, the
JSO is unaffected by the smoothness of the index function. This is
consistent with the way random search algorithms jump from value to
value without taking local regularity into account, as opposed to
gradient-based optimizers for example. Allowing for more iterations also
does not affect success rate significantly. A unit increase in
squintability increases the success rate by 23\%. As dimension increases
by one the success rate almost halves. Increasing the number of jellies
by 10 increases the success rate by 32\%. Table~\ref{tbl-joint-output}
shows the fit statistics supporting that the model is reasonably strong.

Note that the addition of interaction terms for squintability and data
dimension, and smoothness and data dimension, improves the model a small
amount, but does not change the interpretation above. This study intends
to serve as an illustration that the proposed measures effectively
captures the key factors influencing the success rate of PP.

@@

\caption{\label{tbl-mod-output}Jellyfish success rate relative to index
properties and jellyfish hyper-parameters. This is the summary from a
logistic regression fit to smoothness, squintability, dimension, number
of jellyfish and maximum number of iterations. Interestingly,
squintability and dimension strongly affect jellyfish optimization
success. The number of jellies marginally affects success, but index
smoothness, and increasing the number of iterations do not.}

@@

\toprule
\textbf{Characteristic} & \textbf{OR} \textbf{(95\% CI)}\textsuperscript{\textit{1}} & \textbf{p-value} \\ 
\midrule\addlinespace[2.5pt]
Smoothness & 1.01 (0.92 to 1.11) & 0.810 \\ 
Squintability & 1.23 (1.12 to 1.38) & {\bfseries <0.001} \\ 
Dimension & 0.58 (0.45 to 0.73) & {\bfseries <0.001} \\ 
Number of jellyfish & 1.32 (1.17 to 1.50) & {\bfseries <0.001} \\ 
Maximum number of iterations & 1.06 (0.94 to 1.20) & 0.328 \\ 
\bottomrule

@@

\\
\toprule
\textbf{Deviance} & \textbf{DF Residual} & \textbf{Null Deviance} & \textbf{DF Null}\\
\midrule
20.89 & 71 & 49.77 & 76\\
\bottomrule

@@

@@

Using the JSO optimizer for PPGT in the \texttt{tourr} package requires
a change in user behavior. For all the existing optimizers a single
optimization path is followed. The JSO has multiple optimization paths,
and needs additional tools to incorporate into the PPGT, as explained
here.

@@

To use JSO for PPGT in the \texttt{tourr} package, specify
\texttt{search\_f\ =\ search\_jellyfish} in the guided tour.

The animation function \texttt{animate\_*()} won't directly render the
animated projection for JSO, as with other existing optimizers since
multiple tour paths will need to be generated. To visualize the path
visited by each individual jellyfish, assign the animation to an object
(\texttt{res\ \textless{}-\ animate\_xy(...)}). This will save the bases
visited by JSO, along with relevant metadata, as a tibble data object
(see Zhang et al. 2021 for more details on the data object). The bases
visited for individual jellyfish can then be extracted and viewed using
the \texttt{planned\_tour()} function.

@@

The \texttt{ferrn} package (Zhang et al. 2021) has now included
functionality for computing the smoothness and squintability metrics.
Both metrics require sampling random bases using the
\texttt{sample\_bases()} function, followed by the calculation with
\texttt{calc\_smoothness()} or \texttt{calc\_squintability()}.

@@

To sample bases for calculating smoothness, the following inputs are
required: the index function, the dataset, and the number of random
bases to sample. The output of \texttt{sample\_bases()} is a data frame
with a list-column of sampled bases and index values. Parallelization is
available to speed up the index value calculation through the
\texttt{parallel\ =\ TRUE} argument.

The \texttt{calc\_smoothness()} function takes the output from
\texttt{sample\_bases()} and fits a Gaussian process model to the index
values against the sampled bases, as the location, to obtain the
smoothness metric. Starting parameters and additional Gaussian process
arguments can be specified and details of the fit can be accessed
through the \texttt{fit\_res} attribute of the output.

@@

Bases sampling for calculating squintability includes an additional step
of interpolating between the sampled bases and the optimal basis. This
step is performed when the arguments \texttt{step\_size} and
\texttt{min\_proj\_dist} in the \texttt{sample\_bases()} function are
set to non-NA numerical values. Given the projection distance typically
ranging from 0 to 2, it is recommended to set \texttt{step\_size} to 0.1
or lower, and \texttt{min\_proj\_dist} to be at least 0.5 to ensure a
meaningful interpolation length.

The \texttt{calc\_squintability()} function computes squintability using
two methods: 1) parametrically, by fitting a scaled sigmoid function
through non-linear least square (\texttt{method\ =\ "nls"}), and 2)
non-parametrically, using kernel smoothing (\texttt{method\ =\ "ks"}). A
\texttt{bin\_width} argument is required to average the index values
over the projection distance before the fitting. For the parametric
case, the output provides the estimated parameters for the logistic
function (\(\theta_1\) to \(\theta_4\)) and the calculated squintability
metric as Equation (\ref{eq-squintability-parametric}). For the
non-parametric case, it shows the maximum gradient attained
(\texttt{max\_d}), the corresponding projection distance
(\texttt{max\_dist}), and the squintability metric as their products.
